---
title: "HulC on VD"
output:
  html_document:
    code_folding: show
    highlight: haddock
    number_sections: yes
    theme: lumen
    toc: yes
    toc_depth: 4
    toc_float: yes
    df_print: paged
  pdf_document:
    number_sections: yes
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = T, eval = T, warning = F, message=F)
pacman::p_load(dplyr, grf, sandwich, skimr, stargazer,
               lmtest, SuperLearner, ranger, gbm, xgboost, glmnet,
               KernelKnn)
source("utilities.R")

output_type <- knitr::opts_knit$get("rmarkdown.pandoc.to")
```

\pagebreak

This file to reproduce the data analysis on the First Steps program of V\&D (2021) and to compare the confidence interval by their approach and HulC.


```{r}
# copying from VD's code
kc <- read.table("KingCounty2001.data",header=F)
# head(kc)
colnames(kc) <- c("gender","plural","age","race","parity","married","bwt","smokeN","drinkN","firstep","welfare","smoker","drinker","wpre","wgain","education","gestation")
kc$paritycat <- ifelse(kc$parity>1,2,kc$parity)

# set race and paritycat as factors
cols_to_factor <- c("race", "paritycat", "smoker", "gender", "married")
# cols_to_factor <- c("race", "paritycat")
kc[cols_to_factor] <- lapply(kc[cols_to_factor], factor)

# low weight
kc$lbw <- ifelse(kc$bwt<2500,1,0)
```

```{r hyperparam}
alpha <- .05
ci_level <- 1-alpha
SL.library <- c("SL.glm","SL.ranger",
                # SL.svm does not work
                # "SL.svm", 
                "SL.gam",
                "SL.glm.interaction","SL.kernelKnn",
                paste("SL.gam.", 3:10, sep = ""))
```


# EDA

Variable Name | Variable | Description
---------|---------|-------------------
bwt | Infant birth weight (in grams) | Response variable
firstep | first\_step | First steps program participation
age | maternal age | 
gender | child’s sex |
race | race  | asian, black, hispanic, white or other
smoker | smoking status |
paritycat | number of previous live born infants
married | marital status |
wpre | weight prior to pregnancy |
education | education year | 

```{r}
kc %>% 
  select(bwt, firstep, age, gender, race, paritycat, married, smoker, wpre, education) %>%
  skim()
```

# Linear model

The authors consider a linear model with and without interaction between the participation on the first step program and maternal age.

## Full model

```{r, results='asis'}
# Y: infant birth weight (in grams)
# X: participation onthe First Steps program, maternal age, child’s sex, mother’s age, race (asian, black, hispanic, white or other), smoking status and marital status
formula_lm <- bwt ~ firstep+age+gender+race+paritycat+married+smoker+wpre+education
# lm
fit_lm <- lm(formula_lm, kc)
ci_lm <- confint(fit_lm, level = ci_level)

# HulC
ci_hulc <- HulC(kc, FUN = lm, formula = formula_lm)

stargazer(fit_lm, fit_lm,
          type = output_type,
          star.cutoffs = NA,
          column.labels = c("lm", "HulC"),
          ci.custom = list(ci_lm, ci_hulc[,2:3]),
          ci=T,
          keep=c("firstep", "age", "firstep:age"),
          keep.stat = c("n"),
          omit.table.layout = "n")
```

## Age

```{r, results='asis'}
formula_lm <- bwt ~ age+gender+race+paritycat+married+smoker+wpre+education
# lm
fit_lm_age <- lm(formula_lm, kc)
ci_lm_age <- confint(fit_lm_age, level = ci_level)

# HulC
ci_hulc_age <- HulC(kc, FUN = lm, formula = formula_lm)


stargazer(fit_lm_age, fit_lm_age,
          type = output_type,
          star.cutoffs = NA,
          column.labels = c("lm", "HulC"),
          ci.custom = list(ci_lm_age, ci_hulc_age[,2:3]),
          ci=T,
          keep=c("firstep", "age", "firstep:age"),
          keep.stat = c("n"),
          omit.table.layout = "n")
```

## Unadjusted for covariates

```{r, results='asis'}
formula_lm <- bwt ~ firstep
# lm
fit_lm_unadjusted <- lm(formula_lm, kc)
ci_lm_unadjusted <- confint(fit_lm_unadjusted, level = ci_level)

# HulC
ci_hulc_unadjusted <- HulC(kc, FUN = lm, formula = formula_lm)

stargazer(fit_lm_unadjusted, fit_lm_unadjusted,
          type = output_type,
          star.cutoffs = NA,
          column.labels = c("lm (unadjusted)", "HulC (unadjusted)"),
          ci.custom = list(ci_lm_unadjusted, ci_hulc_unadjusted[,2:3]),
          ci=T,
          keep=c("firstep", "age", "firstep:age"),
          keep.stat = c("n"),
          omit.table.layout = "n")
```


## Interaction

```{r, results='asis'}
formula_int <- bwt~firstep+age+firstep*age+gender+race+paritycat+married+smoker+wpre+education

vars <- c('firstep', 'age', 'firstep:age')
# lm with interaction
fit_lm_int <- lm(formula_int, kc)
ci_lm_int <- confint(fit_lm_int, level = ci_level)

# HulC
ci_hulc_int <- HulC(kc, FUN = lm, formula = formula_int)

stargazer(fit_lm_int, fit_lm_int,
          type = output_type,
          star.cutoffs = NA,
          column.labels = c("lm", "HulC"),
          ci.custom = list(ci_lm_int, ci_hulc_int[,2:3]),
          ci=T,
          keep=c("firstep", "age", "firstep:age"),
          keep.stat = c("n"),
          omit.table.layout = "n")
```

## Summary

```{r, results='asis'}
stargazer(fit_lm, fit_lm, fit_lm_age, fit_lm_age,
          # fit_lm_unadjusted, fit_lm_unadjusted,
          fit_lm_int, fit_lm_int,
          type = output_type, 
          star.cutoffs = NA,
          column.labels = c("lm", "HulC", "lm", "HulC", 
                            # "lm (unadjusted)", "HulC (unadjusted)", 
                            "lm", "HulC"),
          ci.custom = list(ci_lm, ci_hulc[,2:3],
                           ci_lm_age, ci_hulc_age[,2:3],
                           # ci_lm_unadjusted, ci_hulc_unadjusted[,2:3],
                           ci_lm_int, ci_hulc_int[,2:3]),
          ci=T,
          dep.var.labels = "Birth weight",
          keep=c("firstep", "age", "firstep:age"),
          keep.stat = c("n"),
          omit.table.layout = "n",
          column.sep.width = "-10pt", digits = 1,
          caption = "Compare the confidence intervals by lm() and HulC. Model (1) and (2) include the First Steps program participation and maternal age; Model (3) and (4) include only age; Model (5) and (6) include the First Steps program participation and maternal age with their interaction. All models control for child’s sex, number of previous live born infants, weight prior to pregnancy, education, smoking status and marital status.",
          label = "tbl:lm-ci",
          out = "output/lm.tex")
```

# GLM

Similarly, the author repeated the analysis after dichotomising the outcome (an infant was considered to have low birth weight if they weighed < 2,500g).

## Full model

```{r, results='asis'}
formula_glm <- lbw~firstep+age+gender+race+paritycat+married+smoker+wpre+education
# glm
fit_glm <- glm(formula_glm, kc, family = "binomial")
ci_glm <- confint(fit_glm, level = ci_level)

# HulC
ci_glm_hulc <- HulC(kc, FUN = glm, formula = formula_glm, family = "binomial")


stargazer(fit_glm, fit_glm,
          type = output_type,
          star.cutoffs = NA,
          column.labels = c("lm", "HulC"),
          ci.custom = list(ci_glm, ci_glm_hulc[,2:3]),
          ci=T,
          keep=c("firstep", "age", "firstep:age"),
          keep.stat = c("n"),
          omit.table.layout = "n")
```

## Age

```{r, results='asis'}
formula_glm <- lbw ~ age+gender+race+paritycat+married+smoker+wpre+education
# lm
fit_glm_age <- glm(formula_glm, kc, family = "binomial")
ci_glm_age <- confint(fit_glm, level = ci_level)

# HulC
ci_glm_hulc_age <- HulC(kc, FUN = glm, formula = formula_glm, family = "binomial") 

stargazer(fit_glm_age, fit_glm_age,
          type = output_type,
          star.cutoffs = NA,
          column.labels = c("lm", "HulC"),
          ci.custom = list(ci_glm_age, ci_glm_hulc_age[,2:3]),
          ci=T,
          keep=c("firstep", "age", "firstep:age"),
          keep.stat = c("n"))
```

## Unadjusted for covariates

```{r, results='asis'}
formula_glm <- lbw ~ firstep
# lm
fit_glm_unadjusted <- glm(formula_glm, kc, family = "binomial")
ci_glm_unadjusted <- confint(fit_glm, level = ci_level)

# HulC
ci_glm_hulc_unadjusted <- HulC(kc, FUN = glm, formula = formula_glm, family = "binomial")

stargazer(fit_glm_unadjusted, fit_glm_unadjusted,
          type = output_type,
          star.cutoffs = NA,
          column.labels = c("lm (unadjusted)", "HulC (unadjusted)"),
          ci.custom = list(ci_glm_unadjusted, ci_glm_hulc_unadjusted[,2:3]),
          ci=T,
          keep=c("firstep", "age", "firstep:age"),
          keep.stat = c("n"))
```


## Interaction 

```{r, results='asis'}
formula_int <- lbw~firstep+age+firstep*age+gender+race+paritycat+married+smoker+wpre+education

# lm with interaction
fit_glm_int <- glm(formula_int, kc, family = "binomial")
ci_glm_int <- confint(fit_glm, level = ci_level)

# HulC
ci_glm_hulc_int <- HulC(kc, FUN = glm, formula = formula_int, family = "binomial") 

stargazer(fit_glm_int, fit_glm_int,
          type = output_type,
          star.cutoffs = NA,
          column.labels = c("lm", "HulC"),
          ci.custom = list(ci_glm_int, ci_glm_hulc_int[,2:3]),
          ci=T,
          keep=c("firstep", "age", "firstep:age"),
          keep.stat = c("n"),
          omit.table.layout = "n")
```
## Summary

```{r, results='asis'}
stargazer(fit_glm, fit_glm, fit_glm_age, fit_glm_age,
          # fit_glm_unadjusted, fit_glm_unadjusted,
          fit_glm_int, fit_glm_int,
          type = output_type, 
          star.cutoffs = NA,
          column.labels = c("lm", "HulC", "lm", "HulC", 
                            # "lm (unadjusted)", "HulC (unadjusted)",
                            "lm", "HulC"),
          ci.custom = list(ci_glm, ci_glm_hulc[,2:3],
                           ci_glm_age, ci_glm_hulc_age[,2:3],
                           # ci_glm_unadjusted, ci_glm_hulc_unadjusted[,2:3],
                           ci_glm_int, ci_glm_hulc_int[,2:3]),
          ci=T,
          dep.var.labels = "Low birth weight",
          keep=c("firstep", "age", "firstep:age"),
          keep.stat = c("n"),
          omit.table.layout = "n",
          column.sep.width = "-10pt", digits = 1,
          caption = "Compare the confidence intervals by glm() and HulC. Model (1) and (2) include the First Steps program participation and maternal age; Model (3) and (4) include only age; Model (5) and (6) include the First Steps program participation and maternal age with their interaction. All models control for child’s sex, number of previous live born infants, weight prior to pregnancy, education, smoking status and marital status.",
          label = "tbl:glm-ci",
          out = "output/glm.tex")
```

# Nonparametric

Assume
\[g\{E(Y|A,L)\} = \beta A + \omega(L)\], 
the authors tried the "partialling out" estimator of Robinson (1988) and their proposal. All the nuisance are estimated using the `SuperLearner` package. 


## "partialling out" estimator of Robinson (1988)

\[\frac{\sum_{i=1}^n \{A_i - \hat{E}(A_i | L_i) \}\{Y_i - \hat{E}(Y_i |L_i) \} }{\sum_{i=1}^n \{A_i - \hat{E}(A_i | L_i) \}^2}\]
where the nuisances are estimated using `SuperLearner`.

### Full model

```{r}
formula <- bwt ~ age+gender+race+paritycat+married+smoker+wpre+education

robinson_fit <- partialling_out_estimator(kc, formula = formula, 
                          Y_var = "bwt", A_var = "firstep",
                          SL.library = SL.library)

robinson_hulc <- HulC(kc, FUN = partialling_out_estimator,
     formula = formula, Y_var = "bwt", A_var = "firstep",
     SL.library = SL.library)
```


```{r, results='asis'}
# output
names(robinson_fit$coefficients) <- "firstep"
fit_template <- lm(bwt~firstep-1, kc)
stargazer(fit_template, fit_template,
          type = output_type,
          coef = list(robinson_fit$coefficients, robinson_fit$coefficients),
          star.cutoffs = NA,
          column.labels = c("lm", "HulC"),
          ci.custom = list(robinson_fit$interval, robinson_hulc[,2:3]),
          ci=T,
          keep=c("firstep"),
          keep.stat = c("n"),
          omit.table.layout = "n")
```

### Age

```{r}
formula <- bwt ~ gender+race+paritycat+married+smoker+wpre+education

robinson_fit_age <- partialling_out_estimator(kc, formula = formula, 
                          Y_var = "bwt", A_var = "age",
                          SL.library = SL.library)

robinson_hulc_age <- HulC(kc, FUN = partialling_out_estimator,
     formula = formula, Y_var = "bwt", A_var = "age",
     SL.library = SL.library)
```

```{r, results='asis'}
# output
names(robinson_fit_age$coefficients) <- "age"
fit_template_age <- lm(bwt~age-1, kc)
stargazer(fit_template_age, fit_template_age,
          type = output_type,
          coef = list(robinson_fit_age$coefficients, robinson_fit_age$coefficients),
          star.cutoffs = NA,
          column.labels = c("lm", "HulC"),
          ci.custom = list(robinson_fit_age$interval, robinson_hulc_age[,2:3]),
          ci=T,
          keep=c("age"),
          keep.stat = c("n"),
          omit.table.layout = "n")
```

### Summary

```{r, results='asis'}
stargazer(fit_template, fit_template,
          fit_template_age, fit_template_age,
          type = output_type,
          coef = list(robinson_fit$coefficients, robinson_fit$coefficients,
                      robinson_fit_age$coefficients, robinson_fit_age$coefficients),
          star.cutoffs = NA,
          column.labels = c("lm", "HulC", "lm", "HulC"),
          ci.custom = list(robinson_fit$interval, robinson_hulc[,2:3],
                           robinson_fit_age$interval, robinson_hulc_age[,2:3]),
          ci=T,
          dep.var.labels = "Birth weight",
          keep=c("firstep", "age", "firstep:age"),
          keep.stat = c("n"),
          omit.table.layout = "n",
          column.sep.width = "-10pt", digits = 1,
          caption = "Compare the confidence intervals by the sandwich package and HulC. The effects are estiamted using the `partialling out' estimator of Robinson. Model (1) and (2) estimate the propensity-overlap weighted effect of First Steps participation on birth weight; Model (3) and (4) estimate the weighted effect of maternal age. All the nuisance functionals are estimated using SuperLearner with all the candidate methods mentioned in the paper except for support vector machines. The estimates are sensitive to the tuning parameters and the candidate methods for the nuisance functional estimation.",
          label = "tbl:robinson-ci",
          out = "output/robinson.tex")
```


## VD proposal


Copying from Page 15

1. Obtain the estimates $\hat{E}(A|L)$ and $\hat{E}(Y|A,L)$, e.g. using machine learning.
2. If $A$ is binary, estimate $E[g\{E(Y |A, L)\}|L]$ as
\[\hat{E}[g\{\hat{E}(Y |A, L)\}|L] = g\{\hat{E}(Y |A = 1, L)\}\hat{E}(A|L) + g\{\hat{E}(Y |A = 0, L)\}\{1 - \hat{E}(A|L)\}\]
otherwise, use an additional machine learning fit (with $g\{\hat{E}(Y |A, L)\}$ as outcome).
3. Obtain an estimate of $\mu(Y, A, L)$:
\[\hat{\mu}(Y, A, L) = g^{-1}\{\hat{E}(Y |A, L)\}\{Y - \hat{E}(Y |A, L)\} +g\{\hat{E}(Y |A, L)\} - \hat{E}[g\{\hat{E}(Y |A, L)\}|L].\]
4. Fit a linear regression of $\mu(Y, A, L)$ on the sole predictor $A - \hat{E}(A|L)$ (without an intercept) using OLS in order to obtain an estimate $\hat{\beta}$ of $\beta$.

The variance is estimated through the sandwich estimator.

### Full model 

```{r}
formula <- lbw ~ age+gender+race+paritycat+married+smoker+wpre+education

VD_fit <- VD(kc, formula = formula, 
   Y_var = "lbw", A_var = "firstep",
   SL.library = SL.library)

VD_hulc <- HulC(kc, FUN = VD, 
     formula = formula, 
     Y_var = "lbw", A_var = "firstep",
     SL.library = SL.library)
```

```{r, results='asis'}
# output
names(VD_fit$coefficients) <- "firstep"
fit_template <- lm(lbw~firstep-1, kc)
stargazer(fit_template, fit_template,
          type = output_type,
          coef = list(VD_fit$coefficients, VD_fit$coefficients),
          star.cutoffs = NA,
          column.labels = c("lm", "HulC"),
          ci.custom = list(VD_fit$interval, VD_hulc[,2:3]),
          ci=T,
          keep=c("firstep"),
          keep.stat = c("n"),
          omit.table.layout = "n")
```


### Age

```{r}
formula <- lbw ~ gender+race+paritycat+married+smoker+wpre+education

VD_fit_age <- VD(kc, formula = formula, 
   Y_var = "lbw", A_var = "age",
   SL.library = SL.library)

VD_hulc_age <- HulC(kc, FUN = VD,
     formula = formula, Y_var = "lbw", A_var = "age",
     SL.library = SL.library)
```

```{r, results='asis'}
# output
names(VD_fit_age$coefficients) <- "age"
fit_template_age <- lm(lbw~age-1, kc)
stargazer(fit_template_age, fit_template_age,
          type = output_type,
          coef = list(VD_fit_age$coefficients, VD_fit_age$coefficients),
          star.cutoffs = NA,
          column.labels = c("lm", "HulC"),
          ci.custom = list(VD_fit_age$interval, VD_hulc_age[,2:3]),
          ci=T,
          keep=c("age"),
          keep.stat = c("n"),
          omit.table.layout = "n")
```

### Summary

```{r, results='asis'}
stargazer(fit_template, fit_template,
          fit_template_age, fit_template_age,
          type = output_type,
          coef = list(VD_fit$coefficients, VD_fit$coefficients,
                      VD_fit_age$coefficients, VD_fit_age$coefficients),
          star.cutoffs = NA,
          column.labels = c("lm", "HulC", "lm", "HulC"),
          ci.custom = list(VD_fit$interval, VD_hulc[,2:3],
                           VD_fit_age$interval, VD_hulc_age[,2:3]),
          ci=T,
          dep.var.labels = "Low birth weight",
          keep=c("firstep", "age", "firstep:age"),
          keep.stat = c("n"),
          omit.table.layout = "n",
          column.sep.width = "-10pt", digits = 1,
          caption = "Compare the confidence intervals by the sandwich package and HulC. The effects are estiamted using the authors' proposals in Section 5.1.2. Model (1) and (2) estimate propensity-overlap weighted effect of First Steps participation on birth weight; Model (3) and (4) estimate the weighted effect of maternal age. All the nuisance functionals are estimated using SuperLearner with all the candidate methods mentioned in the paper except for support vector machines. The estimates are sensitive to the tuning parameters and the candidate methods for the nuisance functional estimation.",
          label = "tbl:VD-ci",
          out = "output/VD.tex")
```